SMT-Boosted Security Types for Low-Level MPC

### 2025-04-24 esop25/slides.tex
checkpoint

\input{slide-macros}

% Title page details: 
\title{Language-Based Security for Low-Level MPC}

@@

\date{September 10, 2024}

@@

  \cemph{Secure Multiparty Computation (MPC)} is a paradigm for
  \cemph{confidential computing}.
  \begin{itemize}
  \item Guarantees data privacy.
  \item Software-based protocols (vs., e.g., enclaves).
  \item No trusted 3rd party.
  \end{itemize}
  Many existing deployments: voting, auctions, satellite trajectories, COVID exposure notification, ...\footnote{https://mpc.cs.berkeley.edu}

@@

  Existing HLLs rely on low-level MPC protocols with correctness
  assumed.
  \begin{itemize}
  \item Low-level protocols possess distinct features, e.g., probabilistic sampling.
  \item Existing proof techniques manual, error prone, complex security model.
  \end{itemize}
  Our contributions:
  \begin{itemize}
  \item Language design for low-level protocol definitions.
  \item (Re)formulation of security model as \cemph{hyperproperties} (e.g., noninterference).
  \item Semi-automated proof techniques.
    \begin{itemize}
    \item Automatically verify small program components, manually combine for whole-program proofs. 
    \end{itemize}
  \end{itemize}

@@

\begin{frame}{The Ideal Functionality}
  A \emph{protocol} $\prog$ implements an \emph{ideal functionality} $\idealf$,
  defined in terms of operations $(+,-,*)$ in a prime field $\mathbb{F}_p$.
  \begin{itemize}
  \item The inputs are \cemph{secret values} of \cemph{clients} in a \emph{federation}
    $\setit{1,...,j}$.
  \item The \cemph{output} is made \emph{publicly available} by $\prog$
  \end{itemize}
  \cemph{The output may leak secret information to an adversary} trying to
  guess secret values.

  \begin{exampleblock}{Consider:}
  $$
  \idealf(x,y) = x * y
  $$
  If the output is odd, the adversary knows both $x$ and $y$ must be, even
  if $x$ and $y$ remain hidden.
  \end{exampleblock}

@@

  Low-level, probabilistic protocols are defined in terms of basic variable types
  (identifiers $x$ distinguish them):
  \begin{itemize}
  \item $\secret{x}$: secret values in $\mathbb{F}_p$.
  \item $\mesg{x}$: messages received during the protocol.
  \item $\flip{x}$: uniformly random samples from $\mathbb{F}_p$ (implemented with
    a random tape semantics).
  \end{itemize}
  We can localize variables and expressions to \emph{clients}, and any information
  exchanged between clients must be through messaging.
  \begin{exampleblock}{Toy Example:}
    $$
    \begin{array}{rcl@{\qquad}l}
      \mx{x}{1} &:=& \elab{(\secret{x} - \flip{x})}{2}  & \textit{(unicast)}\\
      \mx{y}{*} &:=& \elab{(\flip{y} + \mesg{x})}{1}  & \textit{(broadcast)}
    \end{array}
    $$ 
  \end{exampleblock}
\end{frame}

\begin{frame}{(Passive) Threat Model}
  The adversary wants to guess secrets. In addition to knowing the public
  output, the adversary may also corrupt \emph{any} subset of clients $C$:
  \begin{itemize}
  \item adversary has access to the secrets of clients in $C$.
  \item adversary has access to messages received by members of $C$ during execution of $\prog$,
    aka \cemph{adversarial views}.
  \item adversary cannot break the rules of the protocol\footnote{Malicious security also
  considered separately in our work.}.
  \end{itemize}

  \begin{alertblock}{Bounded Declassification}
    The ideal functionality defines a declassification ``limit'', any protocol $\prog$
    should not reveal any more through adversarial views.
  \end{alertblock}
\end{frame}

  \begin{exampleblock}{Example: 3-Party $\fplus$ in $\mathbb{F}_2$ (where $\fplus$ is $\exor$ and $\ftimes$ is $\wedge$)}
  $$
  \idealf(\sx{1}{1},\sx{2}{2},\sx{3}{3}) =
  \sx{1}{1}\ \fplus\ \sx{2}{2}\ \fplus\ \sx{3}{3}
  $$
  Suppose in a run of $\prog$ the output of $\prog$ is 1 and $C = \{ 3 \}$ and
  the adversary knows $\sx{3}{3} = 0$.
  Therefore the adversary knows that either:
  $$
  \sx{1}{1} = 0 \text{\ and\ } \sx{2}{2} = 1
  $$
  or:
  $$
  \sx{1}{1} = 1 \text{\ and\ } \sx{2}{2} = 0
  $$
  with equal joint probability of .5. \cemph{Adversarial views from the run of $\prog$ should not
    allow the adversary better guesses.}
  \end{exampleblock}

\section{Conditional Probabilistic Noninterference}
\begin{frame}{Conditional Probabilistic Noninterference}

  \begin{itemize}
    \item The set of all protocol runs under all possible secrets and random tapes
      induces a probability distribution of program variables.
    \item Conditioning on the output and corrupt secrets =  declassification
      bounds.
    \item Conditioning on adversarial views = distribution of honest secrets.
  \end{itemize}

  \begin{alertblock}{Conditional Noninterference}
    The probability of honest secrets conditioned on the protocol output
    should not be changed by conditioning on adversarial views.

  \begin{itemize}
  \item Sound for the standard \emph{passive simulator security} model.
  %\item A \emph{hyperproperty} that accommodates traditional PL-based security
  %  mechanisms and theory. 
  \end{itemize}
  \end{alertblock}

@@

  \begin{block}{Reconstructive Sharing}
    $$
    \begin{array}{lll}
      \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
      \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} \\ 
      \elab{\mesg{s2}}{1} &:=& \elab{(\secret{2} \fminus \locflip \fminus \flip{x})}{2} \\ 
      \elab{\mesg{s2}}{3} &:=& \elab{\flip{x}}{2} \\ 
      \elab{\mesg{s3}}{1} &:=& \elab{(\secret{3} \fminus \locflip \fminus \flip{x})}{3} \\ 
      \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3}
    \end{array}
    $$
  \end{block}
  \begin{itemize}
  \item $\elab{\mesg{s1}}{2}, \elab{\mesg{s1}}{3}, \elab{\locflip}{1}$ are the \emph{shares} of $\sx{1}{1}$,
    etc.
  \item \emph{algebraically} $\sx{1}{1} = \elab{\mesg{s1}}{2} + \elab{\mesg{s1}}{3} + \elab{\locflip}{1}$.
  \item \emph{probabilistically} any 2 out of three shares are in a uniform random distribution.
  \end{itemize}

@@

\begin{frame}{Example contd.: Public Broadcast Reveal}

  \begin{itemize}
   \item \emph{algebraically} the sum of all shares = $\sx{1}{1} + \sx{2}{2} + \sx{3}{3}$    
  \end{itemize}

  \begin{block}{Broadcast Sum-of-Shares}
    $$
    \begin{array}{lll}
      \mx{ss1}{*} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
      \mx{ss2}{*} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\
      \mx{ss3}{*} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} 
    \end{array}
    $$
  \end{block}

  \begin{itemize}
  \item Messages $\mx{ss1}$ through $\mx{ss3}$ are \cemph{not probabilistically independent of inputs}, but ...
  \item They \emph{are} \cemph{independent of secrets conditioned on the output}!
  \end{itemize}

@@

\begin{frame}{Example contd.: Output}

  \begin{itemize}
    \item Any $\mx{ss1}{*},\mx{ss2}{*},\mx{ss3}{*}$ such that:
    $$
    \mx{ss1}{*} + \ \mx{ss2}{*} + \ \mx{ss3}{*} = \sx{1}{1} + \sx{2}{2}  + \sx{3}{3}
    $$
    are \cemph{equally likely}.
  \end{itemize}

@@

@@

\section{Models of a Protocol}

\begin{frame}{Computing Protocol Distributions}

  The probability distribution of a protocol $\prog$ (as a pmf) can be precisely computed
  by interpretation as a \cemph{constraint program}.
  \begin{itemize}
  \item $\mathbb{F}_2$ interpretation as a Datalog program-- set of Least Herbrand Models.
    \begin{itemize}
      \item Random tapes represented as fact base.
    \end{itemize}
  \item $\mathbb{F}_p$ interpretation as SMT constraints\footnote{Ozdemir, A., Kremer, G., Tinelli, C., Barrett, C. (2023). \emph{Satisfiability Modulo Finite Fields}. In: Enea, C., Lal, A. (eds) Computer Aided Verification. CAV 2023. }-- set of all models.
  \end{itemize}
  Set of models can be queried to enforce hyperproperties-- but poor scaling.

  \begin{alertblock}{Hypothesis: Composable Properties}
    Compositional properties of protocol components can be independently verified, tractably.
  \end{alertblock}

\section{Compositional Properties of Circuits}

\begin{frame}{Prelude and Component Definitions}

  The Prelude language allows modular definition of Overture protocol components.
  \begin{itemize}
  \item Function definitions, function calls, structured data.
  \item A \emph{staged} language that generates a complete Overture protocol
    as a side effect.
  \end{itemize}

  \begin{exampleblock}{Example: Local Summation}
    $$
    \ttt{sum}(x,y,z,i) \{ \mx{z}{i} := \elab{(\mesg{x} \fplus \mesg{y})}{i} \} 
    $$
  \end{exampleblock}

  In addition to programmer convenience, Prelude supports function-level verification.
  %Prelude supports verification:
  %\begin{itemize}
  %\item Prelude functions define Overture (sub)protocols that can be automatically verified.
  %\item Large protocols can be verified manually by composing function properties. 
  %\end{itemize}

\begin{frame}{Circuits and Gate Paradigms}

  Goldreich-Micali-Wigderson (GMW) protocol implements $\idealf$ as a \cemph{circuit}.
  \begin{itemize}
  \item Composed by combing-- aka ``wiring''-- \cemph{gates}.
  \item Each gate standard definition, supports primitive $\fplus$ and $\ftimes$.
  \item Two input wires, one output wire, values remain encrypted on wire. 
  \end{itemize}

  \begin{exampleblock}{Example: 2-Party GMW in $\mathbb{F}_2$}
    Rep.~invariant: each client in $\setit{1,2}$ holds a share of wire values.

    $\ttt{andgmw}(x,y,z)$:

    \begin{itemize}
    \item Input shares $\mx{x}{1}$, $\mx{x}{2}$ and $\mx{y}{1}$, $\mx{y}{2}$--
      wire value for $x$ is $\mx{x}{1} + \mx{x}{2}$, etc.
    \item Output shares $\mx{z}{1}$, $\mx{z}{2}$ received during gate execution.
      $$
      \mx{z}{1} + \mx{z}{2} = (\mx{x}{1} + \mx{x}{2}) * (\mx{y}{1} + \mx{y}{2})
      $$
    \end{itemize}

  \end{exampleblock}

\begin{frame}

  \begin{exampleblock}{Semi-Automatically Verifying 2-Party GMW}

    Picking dummy $x,y,z$, we can execute $\ttt{andgmw}(x,y,z)$
    to generate a generic ``and'' gate $\prog_{\wedge}$ in Overture.
    \begin{itemize}
    \item Automatically calculate the pmf of $\prog_{\wedge}$.
    \item Pmf can be queried to verify critical properties-- e.g., the
      uniform distribution of output shares under any input conditions. 
    \end{itemize}
    Gate verification serves as a \emph{tactic} for manual circuit-level
    proofs. 
    \begin{itemize}
    \item Leverage logics for reasoning about probabilistic programming: Probabilistic
      Separation Logic, Lilac.
    \end{itemize}

  \end{exampleblock}

>>>>

\usepackage{mathpartir}
\usepackage{comment}

% \input{slide-macros}
\input{macros}

% Title page details: 
\title{SMT-Boosted Security Types for Low-Level MPC}

@@

\date{ESOP, May 6, 2025}

@@

  \cemph{Secure Multiparty Computation (MPC)} is a paradigm for
  \cemph{confidential computing}.
  \begin{itemize}
  \item Supports data privacy.
  \item Uses software-based protocols (vs., e.g., enclaves).
  \item No trusted 3rd party.
  \end{itemize}
  Many existing deployments: voting, auctions, satellite trajectories, COVID exposure notification, ...\footnote{https://mpc.cs.berkeley.edu}

@@

  Existing HLLs rely on low-level MPC protocols with correctness
  assumed.
  \begin{itemize}
  \item Distinct features, e.g., probabilistic sampling.
  \item Existing proof techniques mostly manual.
  \end{itemize}
  Our contributions:
  \begin{itemize}
  \item Language design for low-level protocol definitions.
  \item Automated type analysis for security properties, with correctness based on
    \cemph{hyperproperties}\footnote{Skalka and Near, PPDP 2024}.
  \item Our type analysis leverages an \cemph{SMT finite fields theory}\footnote{Ozdemir et. al, CAV 2023} for:
    \begin{itemize}
    \item \cemph{Correctness}
    \item \cemph{Confidentiality}
    \item \cemph{Integrity}
    \end{itemize}
  \end{itemize}

@@

\begin{frame}{What's an MPC Protocol?}
  A \emph{protocol} $\prog$ implements an \emph{ideal functionality} $\idealf$,
  defined in terms of operations $(+,-,*)$ in a prime field $\mathbb{F}_p$.
  \begin{itemize}
  \item The inputs are \cemph{secret values} of \cemph{clients} in a \emph{federation}
    $\setit{1,...,j}$.
  \item Clients are distributed and communicate via \cemph{message passing}.
  \item Any client can be corrupted by the adversary, who can then see their messages.
  \item The \cemph{output} is made \emph{publicly available} by $\prog$.
  \end{itemize}
  \cemph{The output may leak secret information to an adversary} trying to
  guess secret values, since it depends on inputs.

  \cemph{Confidentiality} means that protocol messages don't enable the adversary
  to make better guesses than the output alone.

  \begin{exampleblock}{Example: 3-Party $\fplus$ in $\mathbb{F}_2$ (where $\fplus$ is $\exor$ and $\ftimes$ is $\wedge$)}
  $$
    \idealf =
    \mathit{secret}@{\mathit{client_1}} + \mathit{secret}@{\mathit{client_2}} + 
      \mathit{secret}@{\mathit{client_3}}
  $$
  Suppose in a run of $\prog$ and $\mathit{client_3}$ is
  corrupted, with $\mathit{secret}@{\mathit{client_3}} = 0$. 
  In case the output is 1, the adversary knows that either:
  $$
  \mathit{secret}@{\mathit{client_1}} = 0 \text{\ and\ } \mathit{secret}@{\mathit{client_2}} = 1
  $$
  or:
  $$
  \mathit{secret}@{\mathit{client_1}} = 1 \text{\ and\ } \mathit{secret}@{\mathit{client_2}} = 0
  $$
  with equal joint probability of .5.

  \cemph{Messages sent to $\mathit{client_3}$ during
    a run of $\prog$ should not
    allow the adversary better guesses.}
  \end{exampleblock}

@@

  Low-level, probabilistic protocols are defined in terms of basic variable types
  (identifiers $x$ distinguish them):
  \begin{itemize}
  \item $\secret{x}$: secret values in $\mathbb{F}_p$.
  \item $\mesg{x}$: messages received during the protocol.
  \item $\flip{x}$: uniformly random samples from $\mathbb{F}_p$ (implemented with
    a random tape semantics).
  \end{itemize}
  We localize computation to \emph{clients}, and any information
  exchanged between clients must be through messaging.
  %\begin{exampleblock}{Toy Example:}
  %  $$
  %  \begin{array}{rcl@{\qquad}l}
  %    \mx{x}{1} &:=& \elab{(\secret{x} - \flip{x})}{2}  & \textit{(unicast)}\\
  %    \mx{y}{*} &:=& \elab{(\flip{y} + \mesg{x})}{1}  & \textit{(broadcast)}
  %  \end{array}
  %  $$ 
  %\end{exampleblock}
\end{frame}

@@

  \begin{block}{Reconstructive Sharing}
    $$
    \begin{array}{lll}
      \elab{\mesg{s1}}{2} &:=& \elab{(\secret{1} \fminus \locflip \fminus \flip{x})}{1} \\ 
      \elab{\mesg{s1}}{3} &:=& \elab{\flip{x}}{1} \\ 
      \elab{\mesg{s2}}{1} &:=& \elab{(\secret{2} \fminus \locflip \fminus \flip{x})}{2} \\ 
      \elab{\mesg{s2}}{3} &:=& \elab{\flip{x}}{2} \\ 
      \elab{\mesg{s3}}{1} &:=& \elab{(\secret{3} \fminus \locflip \fminus \flip{x})}{3} \\ 
      \elab{\mesg{s3}}{2} &:=& \elab{\flip{x}}{3}
    \end{array}
    $$
  \end{block}
  \begin{itemize}
  \item $\elab{\mesg{s1}}{2}, \elab{\mesg{s1}}{3}, \elab{\locflip}{1}$ are the \emph{shares} of $\sx{1}{1}$,
    etc.
  \item \emph{algebraically} $\sx{1}{1} \eop \elab{\mesg{s1}}{2} + \elab{\mesg{s1}}{3} + \elab{\locflip}{1}$.
  \item \emph{probabilistically} any 2 out of three shares are in a uniform random distribution.
  \end{itemize}

@@

\begin{frame}{Additive Homomorphism of Shares}

  To sum the secrets \emph{in secret}:

  \begin{block}{Additive Homomorphism: Sum-of-Shares}
    $$
    \begin{array}{lll}
      \mx{ss1}{1} &:=& \elab{(\locflip \fplus \mesg{s2} \fplus \mesg{s3})}{1} \\ 
      \mx{ss2}{2} &:=& \elab{(\mesg{s1} \fplus \locflip \fplus \mesg{s3})}{2} \\
      \mx{ss3}{3} &:=& \elab{(\mesg{s1} \fplus \mesg{s2} \fplus \locflip)}{3} 
    \end{array}
    $$
  \end{block}

  Additive homorphism:
  \begin{itemize}
  \item \emph{Algebraically}, $\mx{ss1}{1} + \mx{ss1}{2} + \mx{ss1}{1} \eop \sx{1}{1} + \sx{2}{2} + \sx{3}{3}$ 
  \item \emph{Probabilistically}, protocol step is noninteractive and each of $\mx{ss1}{i}$ remain in a uniform distribution.
  \end{itemize}

@@

\begin{comment}
\begin{frame}{Example contd.: Output}

  \begin{itemize}
    \item Any $\mx{ss1}{*},\mx{ss2}{*},\mx{ss3}{*}$ such that:
    $$
    \mx{ss1}{*} + \ \mx{ss2}{*} + \ \mx{ss3}{*} \eop \sx{1}{1} + \sx{2}{2}  + \sx{3}{3}
    $$
    are \cemph{equally likely}.
  \end{itemize}

@@

\end{frame}
\end{comment}

\section{SMT Boosting: Correctness}

\begin{frame}{Verifying Correctness}

Letting $\prog$ be the above protocol:
$$
\toeq{\prog} \wedge \ttt{out} \eop \mx{ss1}{1} + \mx{ss1}{2} + \mx{ss1}{1} \models\ \ttt{out} \eop \sx{1}{1} \fplus \sx{2}{2} \fplus \sx{3}{3}
$$

@@

\section{SMT Boosting: Confidentiality}

\begin{frame}{Verifying Confidentiality}

$$    
  \inferrule[Encode]
  {\eqs \models \phi \eop \phi' \oplus \rx{w}{\cid} \\
   \oplus \in \{ \fplus,\fminus \}\\
   \eqj{R}{\eqs}{\phi'}{\ty}}
  {\eqj{R;\{ \rx{w}{\cid} \}}{\eqs}{\phi}{\setit{\cty{\rx{w}{\cid}}{\ty}}}}
$$
